### 2019-08-22 diagnostics/prelresults_brlasso.Rmd
Use BRLasso with some preliminary result to present to SN and OR.

---
title: "Preliminära resultat av BRLasso"
author: "Erik Bulow"
date: '2019-08-22'
output: 
  pdf_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
load("../cache/table1.RData")
table1
```

```{r}
knitr::include_graphics("../graphs/brlasso_roc.png")
```

```{r}
knitr::include_graphics("../graphs/brlasso_auc_ci.png")
```

```{r}
knitr::include_graphics("../graphs/brlasso_separation_auc.png")
```

```{r}
load("../cache/brlasso_tbl_auc.RData")
brlasso_tbl_auc
```

```{r}
load("../cache/brlasso_tbl_coefs.RData")
brlasso_tbl_coefs
```