### 2019-10-17 nhsr/nhsr.Rmd
new draft sent to NH, AG, SN and OR.

---
title: "Prediction-modelling from wheat to bread"
author: "Erik Bülow"
institute: "The Swedish Hip Arthroplasty Register"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

background-image: url(https://i.pinimg.com/originals/77/ff/d7/77ffd7a8dbccd96cc579b5554bed07ff.jpg)

---
class: inverse, center, middle

# Get Started

---