### 2020-06-11 reports/figs.Rmd
prepared figure formats. Fix #32. Fix #42.

<!-- FIGURES -->

```{r flowchartshar, fig.cap = "Flowchart depicting inclusion criteria and number of patients from the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/flowchart_shar.ps")
knitr::include_graphics("../graphs/images.png")
```

```{r flowchartnjr, fig.cap = "Flowchart depicting inclusion criteria and number of patients for the external validation cohort from England and Wales."}
# knitr::include_graphics("../graphs/flowchart_njr.ps")
knitr::include_graphics("../graphs/images.png")

```

```{r rocderivation, fig.cap = "Receiver Operation Characteristics (ROC) curves. The area under the ROC curve (AUC) for the reduced model derived on data from the SHAR was very similar to the main model"}
# knitr::include_graphics("../graphs/roc_derivation.eps")
knitr::include_graphics("../graphs/images.png")

```

```{r rocvalidation, fig.cap = "Receiver Operation Characteristics (ROC) curves. There was no difference when applying the reduced model to the external validation cohort from England and Wales ,compared to the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/roc_validation.eps")
knitr::include_graphics("../graphs/images.png")
```

```{r aucci, fig.cap = "Area Under the Receiver Operation Characteristics Curve (AUC) as a measure of predictive discriminative ability with 95 % bootstrap confidence intervals. AUC above 0.7 were considered good. The reduced model was similar to the main model, and not statistically significantly inferior when used with external data from the NJR."}

# knitr::include_graphics("../graphs/auc_ci.eps")
knitr::include_graphics("../graphs/images.png")
```

```{r separationhist, fig.cap = "The vast majority of patients survived more than 90 days after THA, and therefore the blue bars dominate the histogram (note the scales). There were very few observations with covariate patterns resulting in death probabilities higher than 5 % (green line). Estimated probabilities above this limit are therefore subject to extrapolation."}
# knitr::include_graphics("../graphs/separation_hist.tiff")
knitr::include_graphics("../graphs/images.png")
```

```{r separationdens, fig.cap = "A normalized density plot reveals that patients who died within 90 days were, on average, estimated to have a higher probability to do so. There were very few observations with covariate patterns resulting in death probabilities higher than 5 % (green line). Estimated probabilities above this limit are therefore subject to extrapolation."}
# knitr::include_graphics("../graphs/separation_dens.tiff")
knitr::include_graphics("../graphs/images.png")

```

```{r calibration, fig.cap = "This figure illustrates calibration between observed proportions and predicted probabilities with 95 % confidence intervals. Deviations above the diagonal line indicated that the model under-estimated the probability of death within 90 days. Such deviation was statistically significant for the external validation for predictions below 1.5 %. In contrast, with higher predicted probabilities, the model over-estimated the observed proportions of deaths."}
# knitr::include_graphics("../graphs/calibration.tiff")
knitr::include_graphics("../graphs/images.png")

```