### 2020-08-25 reports/figs.Rmd
Stat figs. Fix #52.

<!-- FIGURES -->

```{r flowchartshar, fig.cap = "Flowchart depicting inclusion criteria and number of patients from the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/flowchart_shar.ps")
knitr::include_graphics("../graphs/images.png")
```

```{r flowchartnjr, fig.cap = "Flowchart depicting inclusion criteria and number of patients for the external validation cohort from England and Wales."}
# knitr::include_graphics("../graphs/flowchart_njr.ps")
knitr::include_graphics("../graphs/images.png")

```

(ref:br1) ~~Outer steps of the bootstrap ranking procedure. All included patients (A) were re-sampled with replacements 100 times (B). For each bootstrap sample, the inner process (C; as depicted in figure \@ref(fig:br2)) was applied. The final output was a table with all potential predictors ranked by the number of times they got selected (D). Potential predictors selected at least once were included in the main model, and each predictor selected at least 33 out of the 100 times, were included in a reduced model as well.~~

```{r br1, fig.cap = "(ref:br1)", out.width = "\\columnwidth"}
# knitr::include_graphics("../graphs/br1.pdf")
knitr::include_graphics("../graphs/images.png")
```

(ref:br2) ~~Inner steps of the bootstrap ranking procedure. For each bootstrap sample (1), 100 new bootstrap samples were created (2). Logistic LASSO-regression was applied within each sample (3). Absolute values of the estimated coefficient values (whereof some shrunk to 0 by the LASSO), were averaged as a measure of variable importance (4). Variables with their estimated variable importance above an estimated break-point from linear piece-wise regression (5) were kept as potential predictors (6).~~

```{r br2, fig.cap = "(ref:br2)", out.width = "\\columnwidth"}
# knitr::include_graphics("../graphs/br2.pdf")
knitr::include_graphics("../graphs/images.png")
```

```{r rocderivation, fig.cap = "Receiver Operation Characteristics (ROC) curves. The area under the ROC curve (AUC) for the reduced model derived on data from the SHAR was very similar to the main model"}
# knitr::include_graphics("../graphs/roc_derivation.eps")
knitr::include_graphics("../graphs/images.png")

```

```{r rocvalidation, fig.cap = "Receiver Operation Characteristics (ROC) curves. There was no difference when applying the reduced model to the external validation cohort from England and Wales ,compared to the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/roc_validation.eps")
knitr::include_graphics("../graphs/images.png")
```

```{r aucci, fig.cap = "Area Under the Receiver Operation Characteristics Curve (AUC) as a measure of predictive discriminative ability with 95 % bootstrap confidence intervals. AUC above 0.7 were considered good. The reduced model was similar to the main model, and not statistically significantly inferior when used with external data from the NJR."}

# knitr::include_graphics("../graphs/auc_ci.eps")
knitr::include_graphics("../graphs/images.png")
```

```{r separationhist, fig.cap = "The vast majority of patients survived more than 90 days after THA, and therefore the blue bars dominate the histogram (note the scales). There were very few observations with covariate patterns resulting in death probabilities higher than 5 % (green line). Estimated probabilities above this limit are therefore subject to extrapolation."}
# knitr::include_graphics("../graphs/separation_hist.tiff")
knitr::include_graphics("../graphs/images.png")
```

```{r separationdens, fig.cap = "A normalized density plot reveals that patients who died within 90 days were, on average, estimated to have a higher probability to do so. There were very few observations with covariate patterns resulting in death probabilities higher than 5 % (green line). Estimated probabilities above this limit are therefore subject to extrapolation."}
# knitr::include_graphics("../graphs/separation_dens.tiff")
knitr::include_graphics("../graphs/images.png")

```

```{r calibration, fig.cap = "This figure illustrates calibration between observed proportions and predicted probabilities with 95 % confidence intervals. Deviations above the diagonal line indicated that the model under-estimated the probability of death within 90 days. Such deviation was statistically significant for the external validation for predictions below 1.5 %. In contrast, with higher predicted probabilities, the model over-estimated the observed proportions of deaths."}
# knitr::include_graphics("../graphs/calibration.tiff")
knitr::include_graphics("../graphs/images.png")

```