### 2020-09-05 reports/tabs.Rmd

# Table headings

##### PAGE BREAK

(ref:comorbidity) Categorization of comorbidities based on diagnostic groups according to the Charlson and Elixhauser comorbidity indices. ~~Each main  comorbidity category was first defined based on comorbidity data from the Swedish National patient register and from the Hospital Episodes statistics registry in England and Wales, coded along the International Classification of Diseases version 10 (ICD-10) and categorized according to Quan et al.[@Quan2005]~~ (CNS = central nervous system.)

```{r tabcategorization}
tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`),
    `Comorbidities by groups` = 
      gsub("Obesity", "~~Diagnosed~~ obesity", `Comorbidities by groups`)
  ) %>% 
  mutate_all(~ gsub("nemia", "naemia", .)) %>% 
  mutate_all(~ gsub("tumor", "tumour", .)) %>% 
  knitr::kable(
    caption = "(ref:comorbidity)")

```

##### PAGE BREAK

(ref:agedist) ~~Age distribution among patients who died in the Swedish derivation cohort (range `r age_range`years).~~
```{r agedist}
knitr::kable(
  tab_agedist,
  caption = "(ref:agedist)"
)

```

##### PAGE BREAK

```{r tab1}
table1_combined_tot %>% 
  mutate(
    what = gsub("Heart infarct", "Myocardial infarction", what),
    what = gsub("Obesity", "~~Diagnosed~~ obesity", what),
    what = gsub("Anemia", "Anaemia", what)
  ) %>% 
knitr::kable(
    caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anaesthesiologists classification. CNS = central nervous system.")
```

##### PAGE BREAK

```{r brlprop}
t <- 
  brlasso_tbl_selected %>% 
  mutate(
    variable = clean_names(variable),
    variable = gsub("= 2", "II", variable),
    variable = gsub("= 3", "III", variable),
    variable = gsub("Heart infarct", "Myocardial infarction", variable),
    variable = gsub("Obesity", "~~Diagnosed~~ obesity", variable),
    variable = gsub("Anemia", "Anaemia", variable)
  )

knitr::kable(
  t, 
  caption = "Variables selected by the bootstrap ranking procedure, and therefore kept in the main model. Only variables chosen at least 33 times out of 100 were kept in the reduced model. ASA classes II and III were, however, kept in both models. Variables chosen more frequently are likely to be more important predictors. (CNS = central nervous system.)")
```

##### PAGE BREAK

```{r brlreduced}
brlasso_tbl_coefs %>% 
  select(-math) %>%
  mutate(
    term = gsub("= 2", "II", term),
    term = gsub("= 3", "III", term),
    term = gsub("Obesity", "~~Diagnosed~~ obesity", term)
  ) %>% 
knitr::kable(
  caption = 'Estimated coefficients (beta) and odds ratios (OR) with 95 % confidence intervals for the reduced model.'
)
```